---
title: "Lab 03 - Nobel laureates"
author: "Graham Overheu"
date: "1/29/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

There are `nrow(nobel)` observations and `ncol(nobel)` variables. Each row represents one individual Nobel Prize given out to the winner.

### Exercise 2

```{r}
nobel_living <- nobel_living %>% 
  filter(
    is.na(died_date),
    !is.na(country),
    gender != "org"
  )
```
There are `nrow(nobel_living)` observations in the data set.

### Exercise 3
```{r}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

The graph below shows that for every science category, the majority of Nobel Prize winners were based in the USA when they won the award, particularly for the field of economics.

### Exercise 4


### Exercise 5

Yes, the graph supports BuzzFeed's claim that of those US-based Nobel laureates, many of them were born in other countries. While there weren't many compared to the number of those who were born in the US, it can still be said that a reasonable amount of them were not born in the US; for instance, for medicine, the number of laureates who received the award in the USA but were born in a different country takes up around 25-30% of the "USA" bar.

### Exercise 6
Germany and the United Kingdom are the most common.
